The Practical Application of Computerised Multistage Tests

### 2021-02-24 article.Rmd
Added MIFs for all stages

---
title: "The Practical Application of Computerised Multistage Tests"
author:
- name: James Page
  affiliation: SP Educational Consulting
  affiliation_url: https://www.speduconsulting.co.uk/
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document:
    toc: true
    toc_depth: 2
    highlight: haddock
  html_document: default
version: 1
description: |
  A practical guide to the design and development of computerised multistage tests.
bibliography: references.bib
---

@@

```{r IRT helper functions}
# return a data frame with true score summaries
# that can then be plotted
trueScoreSummary <- function(plotData, module=NA) {
  data.frame(module = module, 
             theta = as.vector(plotData[,1]), 
             trueScore = as.vector(rowSums(plotData[,-1])))
}
```

@@

```{r Panel 1, Stage 1 MCC}
moduleA1_items <- which(colnames(dfTrial) %in% colnames(dfModuleAItems))
plotData <- trueScoreSummary(plot(raschmodel, 
                                  type = "ICC", 
                                  items = moduleA1_items, plot = FALSE), 
                             "A")
ggplot(plotData, aes(x=theta, y=trueScore)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="True Score")
```
You can read this off such that a candidate who scores 9 in the module A section of the test would have an estimated ability score, ot $\hat{\theta}$ of approximately 2.

@@

```{r Panel 1, Stage 2 MCCs}
moduleB1_items <- which(colnames(dfTrial) %in% colnames(dfModuleBItems))
moduleC1_items <- which(colnames(dfTrial) %in% colnames(dfModuleCItems))

@@

```{r Panel 1, Stage 3 MCCs}
moduleD1_items <- which(colnames(dfTrial) %in% colnames(dfModuleDItems))
moduleE1_items <- which(colnames(dfTrial) %in% colnames(dfModuleEItems))
moduleF1_items <- which(colnames(dfTrial) %in% colnames(dfModuleFItems))
moduleG1_items <- which(colnames(dfTrial) %in% colnames(dfModuleGItems))

@@

```{r}
itemNo <- 7
itemID <- str_extract(rownames(df_difficulties)[itemNo], "Item\\d+")
itemDifficulty <- df_difficulties[itemNo]
itemPlot <- data.frame(plot(raschmodel, type = "IIC", items = itemNo, plot = FALSE))
colnames(itemPlot) <- c('theta', 'information')
ggplot(itemPlot, aes(x=theta, y=information)) + 
  geom_line() +
  labs(title="Item Information Function", x="θ", y="I(θ)")
```

>>>>

---
title: "The Practical Application of Computerised Multistage Tests"
author:
- name: James Page
  affiliation: SP Educational Consulting
  affiliation_url: https://www.speduconsulting.co.uk/
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    highlight: haddock
    latex_engine: lualatex
  word_document: default
  html_document: default
version: 1
description: |
  A practical guide to the design and development of computerised multistage tests.
bibliography: references.bib
---

@@

```{r IRT helper functions}
# return a data frame with true score summaries
# that can then be plotted
trueScoreSummary <- function(plotData, module=NA) {
  data.frame(module = module, 
             theta = as.vector(plotData[,1]), 
             trueScore = as.vector(rowSums(plotData[,-1])))
}

# return a data frame with information coefficient summaries
# that can then be plotted
informationCoefficients <- function(plotData, module=NA) {
  data.frame(module = module, 
             theta = as.vector(plotData[,1]), 
             information = as.vector(rowSums(plotData[,-1])))
}
```

@@

```{r Stage 1 MCC}
moduleA1_items <- which(colnames(dfTrial) %in% colnames(dfModuleAItems))
plotData <- trueScoreSummary(plot(raschmodel, 
                                  type = "ICC", 
                                  items = moduleA1_items, plot = FALSE), 
                             "A")
ggplot(plotData, aes(x=theta, y=trueScore)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="True Score")
```
You can read this off such that a candidate who scores 9 in the module A section of the test would have an estimated ability score, ot $\hat{\theta}$ of approximately 2.

@@

```{r Stage 2 MCCs}
moduleB1_items <- which(colnames(dfTrial) %in% colnames(dfModuleBItems))
moduleC1_items <- which(colnames(dfTrial) %in% colnames(dfModuleCItems))

@@

```{r Stage 3 MCCs}
moduleD1_items <- which(colnames(dfTrial) %in% colnames(dfModuleDItems))
moduleE1_items <- which(colnames(dfTrial) %in% colnames(dfModuleEItems))
moduleF1_items <- which(colnames(dfTrial) %in% colnames(dfModuleFItems))
moduleG1_items <- which(colnames(dfTrial) %in% colnames(dfModuleGItems))

@@

```{r Item Information Function}
itemNo <- 7
itemID <- str_extract(rownames(df_difficulties)[itemNo], "Item\\d+")
itemDifficulty <- df_difficulties[itemNo]
itemPlot <- data.frame(plot(raschmodel, type = "IIC", items = itemNo, plot = FALSE))
colnames(itemPlot) <- c('theta', 'information')
ggplot(itemPlot, aes(x=theta, y=information)) + 
  geom_line() +
  labs(title="Item Information Function", x="θ", y="I(θ)")
```

### Module Information Functions (MIFs)

Just as with the ICC-MCC relationship, its possible to sum a set of IIFs into a Module Information Function (MIF).

## Routing and Cut-Scores

Routing between modules is determined by the information functions for the modules in the next stage. We should aim to place the cut score to be that at which the MIFs cross.

```{r Stage 1 MIF}
plotData <- informationCoefficients(plot(raschmodel, 
                                         type = "IIC", 
                                         items = moduleA1_items, 
                                         plot = FALSE), 
                                    "A")

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="I(θ)")
```
You can see that the maximum information point Module A is pretty close to $\theta = 0$.

Now lets look at the MIFs for the two stage 2 modules:

```{r Stage 2 MIFs}
plotData <- rbind(
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleB1_items, plot = FALSE), 
                          "B"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleC1_items, plot = FALSE), 
                          "C"))

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="I(θ)") +
  geom_vline(xintercept = 0.81, linetype="dotted")
```
Here we can see that there is point of intersection between modules B and C. This suggests, that candidates who have a $\hat\theta < 0.81$ in Module A should be routed to Module B, and all others routed to Module C.

We can then repeat the exercise for the stage 3 modules.

```{r Stage 3 MIFs}
plotData <- rbind(
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleD1_items, plot = FALSE), 
                          "D"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleE1_items, plot = FALSE), 
                          "E"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleF1_items, plot = FALSE), 
                          "F"))

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="I(θ)")

plotData <- rbind(
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleE1_items, plot = FALSE), 
                          "E"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleF1_items, plot = FALSE), 
                          "F"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleG1_items, plot = FALSE), 
                          "G"))

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Characteristic Curve", x="θ", y="I(θ)")
```
